An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-28 VLDB2026/main.tex
update fig
modify

		\footnotesize{
			\begin{minipage}{\linewidth}
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport.
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{Dynamic} indicates whether the algorithm supports the operation of dynamically inserting data points.
			\end{minipage} 
			\vspace{-0.1cm}
		}

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation. In contrast, CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. Compared with CAPS, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, degenerating into a situation closer to inverted retrieval, and the effect may be greatly impaired.

	\textit{Performance of Algorithms under Multi-Attribute Building (M vs. M).}  
	As shown in Figure \ref{fig:exp_2_1}, IVF-based methods (Puck and CAPS) significantly outperform graph-based methods. The reason for Puck's superior performance is that it filters out irrelevant results during the retrieval process rather than searching for more candidates. The performance of CPAS is as good as that of Puck. The difference between them is that CPAS further groups the results by attributes within clusters. In contrast, FilteredVamana performs the worst because its multi-attribute indexing forces neighbor selection to cover multiple attributes, thereby reducing the efficiency of single-attribute queries. This results in excessive traversal of irrelevant nodes, making it slower than other methods.

@@

	\textit{\textbf{Multi-Attribute Building and Multi-Attribute Query.}}  
	Compared to single-attribute filtering, multi-attribute joint filtering better reflects \textcolor{red}{real-world scenarios}. To evaluate algorithm performance \textcolor{red}{under this setting}, we construct the index using 3 uniformly distributed attributes and \textcolor{red}{apply filtering on all 3 attributes simultaneously during queries}.
%	Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world application scenarios. To evaluate algorithm performance under such scenario, we construct the index using 3 uniformly distributed attributes and apply filtering conditions on all 3 attributes simultaneously during the query process.

@@

	To investigate this effect, we evaluate 4 AS levels (1\%, 25\%, 50\%, and 75\%) on the SIFT1M dataset. Using a single-thread execution environment, we compare the query performance of 13 algorithms under varying AS conditions, as shown in Figure~\ref{fig:exp_5_1_1_SingleLabel_1thread}.

>>>>

		\footnotesize{
			\begin{minipage}{\linewidth}
				\textsuperscript{Y} Support, 
				\textsuperscript{N} Unsupport.
				\textit{Base} indicates that the hybrid query algorithm is an improvement based on a specific ANN search algorithm. 
				\textit{Dynamic Insert} indicates whether the algorithm supports the operation of dynamically inserting data points.
			\textit{Multi Thread} indicates whether the algorithm supports multi-threaded search.
			\end{minipage} 
			\vspace{-0.1cm}
		}

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation.  CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, degenerating into a situation closer to inverted retrieval, and the effect may be greatly impaired.

	\textit{Performance of Algorithms under Multi-Attribute Building (M vs. M).}  
	As shown in Figure \ref{fig:exp_2_1}, IVF-based methods (Puck and CAPS) significantly outperform graph-based methods. The reason for Puck's superior performance is that it filters out irrelevant results during the retrieval process rather than searching for more candidates. The performance of CAPS is as good as that of Puck. The difference between them is that CAPS further groups the results by attributes within clusters. In contrast, FilteredVamana performs the worst because its multi-attribute indexing forces neighbor selection to cover multiple attributes, thereby reducing the efficiency of single-attribute queries. This results in excessive traversal of irrelevant nodes, making it slower than other methods.

@@

	\textit{\textbf{Multi-Attribute Building and Multi-Attribute Query.}}  
	Compared to single-attribute filtering, multi-attribute joint filtering better reflects \textcolor{red}{real-world scenarios}. To evaluate algorithm performance \textcolor{red}{under this setting}, we construct the index using 3 uniformly distributed attributes and \textcolor{red}{apply filtering on all 3 attributes simultaneously during queries}.
	 \textcolor{red}{As Filtered-DiskANN does not support the AND rule used in our multi-attribute queries, it is excluded.}
	%	Compared to single-attribute filtering, multi-attribute joint filtering better reflects real-world application scenarios. To evaluate algorithm performance under such scenario, we construct the index using 3 uniformly distributed attributes and apply filtering conditions on all 3 attributes simultaneously during the query process.

@@

	To investigate this effect, we evaluate 4 AS levels (1\%, 25\%, 50\%, and 75\%) on the SIFT1M dataset. Using a single-thread execution environment, we compare the query performance of attribute filtering algorithms under varying AS conditions, as shown in Figure~\ref{fig:exp_5_1_1_SingleLabel_1thread}.
